Discrete Quantum Theories and Computing

### 2019-01-12 dissertation.tex
dissertation.lyx "Preface" and "Chapter 6 Further Discussion" Updating...

\begin{preface*}
\begin{itemize}
\item The latest version of this thesis is in GitHub: \code{https://git.io/fxbuG}
\item Its source code is typeset using \LyX ~\cite{LyX} in: \code{https://git.io/fxbuE}
\item Any comments can be left in the Issue part of the GitHub Repository:
\code{https://git.io/fxbug}
\end{itemize}
\end{preface*}

@@

\begin{comment}
When people proved the original Gleason theorem, people usually exploited
the geometrical structure of real 3-dimensional Hilbert space \cite{gleason1957,peres1995quantum,RichmanBridges1999,Hamhalter2013}.
Since our finite-precision extension of the Gleason theorem only applies
on a class of QIVPMs, we might want to ask how to modify these geometrical
arguments to have a Gleason-type theorem for general QIVPMs. We will
further study the tensor product structure among QIVPMs which is essential
for defining product and entangled states, and serves the basis to
discuss quantum nonlocality~\cite{Bell1964,Redhead1987-REDINA,peres1995quantum,Jaeger2007}
and quantum computing with QIVPMs. Finally, we want to improve the
discrete quantum theories to consider QIVPMs over finite fields in
future research.
\end{comment}
As we discussed in the previous section, the finite-precision extension
of the Gleason theorem, Thm.~\ref{thm:Finite-precision-Gleason},
only applies on the QIVPMs composted by interval maps $f:\left[0,1\right]\rightarrow\mathscr{I}$
and quantum probability measures $\muB_{\rho}:\events\rightarrow\left[0,1\right]$,
where the maximum length of intervals in $\mathscr{I}$ is an important
. In this case, the states consistent with the composition $f\circ\muB_{\rho}$
is bounded by . While the QIVPM defined in Table~\ref{tab:non-Born-QIVPM}
is not in this class, it is too wild to have a non-empty core, and
the maximum length of its intervals $\mathscr{I}_{0}$ is $1$, the
longest length of all possible intervals. This suggests us to ask
if we shirk the maximum length of intervals alone, whether this ill-behaved
QIVPMs disappears. More precisely, we have the following question.

\begin{question}In a Hilbert space $\Hilb$ of dimension $D\geq3$,
given a QIVPM $\bar{\mu}:\events\rightarrow\mathscr{I}$, $\alpha$
be the maximum length of intervals in $\mathscr{I}$, is there an
$\varepsilon>0$ such that $\coreBorn\left(\bar{\mu},\events\right)\ne\emptyset$
for $\alpha<\varepsilon$.\end{question}

\noindent %
\begin{comment}

@@

\todo{Add a two-page conclusion chapter which could include further
discussion, but it can be only a few sentences? }\todo{Discrete
Schrödinger equation? And its relation with Grover's algorithm? }\todo{Notice
that in differential geometry, Killing–Hopf theorem asserts that having
positive constant curvature guarantees a manifold is essentially a
sphere, but say little things even if the curvature has only small
deviation from constant. Since a quantum probability measure is glued
by classical probability measures, like a manifold. When their relations
are exact, Gleason's theorem asserts that a quantum probability measure
can essentially induced by the Born rule, but they can be wild when
we move on to interval-valued situation! }
\end{comment}

>>>>

\begin{preface*}
\begin{itemize}
\item The latest PDF version of this thesis is in GitHub: \code{https://git.io/fxbuG}.
\item Its source code is typeset using \LyX ~\cite{LyX} in \code{https://git.io/fxbuE}
and the exported \LaTeX{} file is in \code{https://git.io/fhZBW}.
\item Any comments can be left in the Issue part of the GitHub repository:
\code{https://git.io/fxbug}.
\end{itemize}
Following explains how to compile the PDF from the \LaTeX{} source:
\begin{enumerate}
\item \label{enu:Clone-the-git}Clone the git repository of this thesis
\code{https://git.io/fhZP5} into your local directory and the thesis
document class \code{https://github.com/yuttai/iuphd} in a location
where the system can find.
\item In the local directory of this thesis, run \code{xelatex dissertation.tex}.
\item Run \code{biber dissertation}.
\item Run \code{xelatex dissertation.tex} several times until all the references
are resolved.
\end{enumerate}
The previous steps only tested in Microsoft Windows. Please report
me if there is any difficulty to compile it. Besides, to compile a
PDF file from the \LyX{} source, just finish step~\ref{enu:Clone-the-git}
and click the toolbar button \includegraphics[width=1em]{buffer-view}.\end{preface*}

@@

\begin{comment}
When people proved the original Gleason theorem, people usually exploited
the geometrical structure of real 3-dimensional Hilbert space \cite{gleason1957,peres1995quantum,RichmanBridges1999,Hamhalter2013}.
Since our finite-precision extension of the Gleason theorem only applies
on a class of QIVPMs, we might want to ask how to modify these geometrical
arguments to have a Gleason-type theorem for general QIVPMs. We will
further study the tensor product structure among QIVPMs which is essential
for defining product and entangled states, and serves the basis to
discuss quantum nonlocality~\cite{Bell1964,Redhead1987-REDINA,peres1995quantum,Jaeger2007}
and quantum computing with QIVPMs. Finally, we want to improve the
discrete quantum theories to consider QIVPMs over finite fields in
future research.
\end{comment}
As we discussed in the previous section, the finite-precision extension
of the Gleason theorem, Thm.~\ref{thm:Finite-precision-Gleason},
only applies on the QIVPMs composed by interval maps $f:\left[0,1\right]\rightarrow\mathscr{I}$
and quantum probability measures $\muB_{\rho}:\events\rightarrow\left[0,1\right]$:
the states consistent with the composition $f\circ\muB_{\rho}$ converges
to $\rho$ as the maximum length of intervals in $\mathscr{I}$ shrinking
to $0$. In contrast, the wildest QIVPM defined in Table~\ref{tab:non-Born-QIVPM}
maps to $\mathscr{I}_{0}$ whose intervals' maximum length is $1$,
the longest possible length. This provides another possibility that
shirking the maximum length of intervals itself could result the QIVPMs
mapping to these intervals having non-empty cores. More precisely,
we have the following question.

\begin{question}\label{question:approximation-Gleason}Given a Hilbert
space $\Hilb$ of dimension $D\geq3$, is there an $\varepsilon>0$
such that for all QIVPM $\bar{\mu}:\events\rightarrow\mathscr{I}$
satisfying the maximum length of intervals in $\mathscr{I}$ less
than $\varepsilon$, $\bar{\mu}$ must have an non-empty core, i.e.,
$\coreBorn\left(\bar{\mu},\events\right)\ne\emptyset$.\end{question}

@@

\todo{Add a two-page conclusion chapter which could include further
discussion, but it can be only a few sentences? }\todo{Discrete
Schrödinger equation? And its relation with Grover's algorithm? }\todo{Notice
that in differential geometry, Killing–Hopf theorem asserts that having
positive constant curvature guarantees a manifold is essentially a
sphere, but say little things even if the curvature has only small
deviation from constant. Since a quantum probability measure is glued
by classical probability measures, like a manifold. When their relations
are exact, Gleason's theorem asserts that a quantum probability measure
can essentially induced by the Born rule, but they can be wild when
we move on to interval-valued situation! }
To better understand this question, consider the $D=3$ situation
which is largely simplified by the fact that any two-dimensional projectors
can be expressed as the complement of a one-dimensional projector,
and by Eq.~(\ref{eq:QIVPM-complement}) so does their interval-valued
probabilities, i.e., $\bar{\mu}\left(\mathbb{1}-\proj{\psi}\right)=\necess-\bar{\mu}\left(\proj{\psi}\right)$.
Hence, a QIVPM is completely determined by its values on the one-dimensional
projectors which are one-to-one corresponding to the irreducible states,
and these irreducible states are encoded in the complex projective
space $\CP{2}$ as discussed in Sec.~\ref{subsec:Explicit-generalization-of}.
To further simplify the question, for any state $\ket{\psi}\in\CP{2}$,
we can decompose interval-valued probabilities on these states into
its left-end and right-end, where $\left[\frameL\left(\ket{\psi}\right),\frameR\left(\ket{\psi}\right)\right]=\bar{\mu}\left(\proj{\psi}\right)$.
According to Lemma~\ref{thm:convex-3-1}, $\bar{\mu}:\events\rightarrow\mathscr{I}$
is a QIVPM if and only if the pair of functions $\frameL:\CP{2}\rightarrow\left[0,1\right]$
and $\frameR:\CP{2}\rightarrow\left[0,1\right]$ satisfying the following
interval-inclusion
\left[1,1\right]-\left[\frameL\left(\ket{\psi_{0}}\right),\frameR\left(\ket{\psi_{0}}\right)\right]\subseteq\left[\frameL\left(\ket{\psi_{1}}\right),\frameR\left(\ket{\psi_{1}}\right)\right]+\left[\frameL\left(\ket{\psi_{2}}\right),\frameR\left(\ket{\psi_{2}}\right)\right]\label{eq:interval-valued-frame-function}
for any orthonormal basis $\left\{ \ket{\psi_{i}}\right\} _{i=0}^{2}$.
This interval-inclusion can be rephrased as a long inequality
\frameL\left(\ket{\psi_{1}}\right)+\frameL\left(\ket{\psi_{2}}\right)\le1-\frameR\left(\ket{\psi_{0}}\right)\le1-\frameL\left(\ket{\psi_{0}}\right)\le\frameR\left(\ket{\psi_{1}}\right)+\frameR\left(\ket{\psi_{2}}\right)\,.
Since the maximum length of intervals in $\mathscr{I}$ is less than
$\varepsilon$, $\left[\frameL\left(\ket{\psi}\right),\frameR\left(\ket{\psi}\right)\right]\in\mathscr{I}$
implies $\frameR\left(\ket{\psi}\right)-\frameL\left(\ket{\psi}\right)<\varepsilon$.
Then, the largest term in the previous inequality $\frameR\left(\ket{\psi_{1}}\right)+\frameR\left(\ket{\psi_{2}}\right)$
is bounded by $\frameL\left(\ket{\psi_{1}}\right)+\frameL\left(\ket{\psi_{2}}\right)+2\varepsilon$.
In another word, the left-end function satisfies the following inequalities
\begin{eqnarray}
 &  & \frameL\left(\ket{\psi_{1}}\right)+\frameL\left(\ket{\psi_{2}}\right)\le1-\frameL\left(\ket{\psi_{0}}\right)\le\frameL\left(\ket{\psi_{1}}\right)+\frameL\left(\ket{\psi_{2}}\right)+2\varepsilon\,.\\
 & \Leftrightarrow & 1-2\varepsilon\le\sum_{i=0}^{2}\frameL\left(\ket{\psi_{i}}\right)\le1\,.\label{eq:left-end-frame-function}
\end{eqnarray}
In this language, Gleason's theorem basically states that when $\varepsilon=0$,
given any function $\frameL:\CP{2}\rightarrow\left[0,1\right]$ satisfying
Eq.~(\ref{eq:left-end-frame-function}), there exists a unique mixed
state $\rho=\sum_{j=1}^{N}q_{j}\proj{\Phi_{j}}$ such that 
\frameL\left(\ket{\psi}\right)=\melem{\psi}{\rho}{\psi}\label{eq:Gleason-for-CP2}
for any $D$-dimensional projector~$P$, where $\ket{\Phi_{j}}\in\mathcal{H}$
are normalized, $q_{j}>0$, and $\sum_{j=1}^{N}q_{j}=1$. Our Question~\ref{question:approximation-Gleason}
ask what $\frameL$ would look like with a positive $\varepsilon$.

With different settings, whether there is an approximate version of
Gleason's theorem in general was asked by Sam Sanders in constructivenews~\cite{Sanders2013},
and there is no clear answer for his question as far as I understand.
To have an idea how hard this question could be, recall in Sec.~\ref{sec:fuzzy}
we state that a quantum probability space is glued by a family of
classical probability spaces. This is similar to the situation that
a manifold is glued by many local coordinates. When the curvature
of each small piece is ``exactly'' the same and positive, Killing-Hopf
theorem asserts this manifold is a sphere, but little can we say even
if the curvature has only small deviation from constant. Similar situation
might happen when approximating Gleason's theorem. Since the sum of
$\frameL$ on each local classical probability space is ``exactly''
the same and equal to $1$, Gleason's theorem can assert $\frameL$
can be expressed as Eq.~\ref{eq:Gleason-for-CP2}. However, when
each local classical probability space becomes imprecise, a general
$\frameL$ might be as wild as we can image, and we might say little
about its global property.
